---
title: "Classification"
author: Darius Goergen
site: workflowr::wflow_site
output:
  workflowr::wflow_html:
    toc: true
editor_options:
  chunk_output_type: console
bibliography: library.bib
csl: elsevier-harvard.csl
link-citations: yes
---

```{r calibration-setup, echo=FALSE, warning=FALSE, message=FALSE}
source("code/setup_website.R")
source("code/functions.R")
data = read.csv(file = paste0(ref, "reference_database.csv"), header = TRUE)
```

## Overview

The overall aim of this project was to ease the process of classifying the spectra
of particles found in environmental samples. A script was created which
takes `.txt` files in the `smp` directory and classifies them based on the decision
fusion explained [here](calibration.html).
The script expects one file for each sample. The names of the files need to be unique
and will be used as an identifier in the output. It is also expected that the file
consists of two columns. The first column indicates the wavenumbers of the sample
as a numeric, while the second column contains information about the reflectance value.

## Preparing Classification

The head of the classification script contains some variables that can be changed.
This is important if extensions were made to the database or a new calibration was to be done.
In this case, the directory containing the models can be specified with the `MODEL`
variable. For now, however, the variables hint towards the `BASE` models which were created
during the first calibration. Also, the type of the classification can be specified in the
`TYPE` variable. Currently, the fusion of all four models is chosen, but any single model
can be selected. Finally, the `FORMAT` variable indicates the file extension of the sample
data, since it is also possible to provide spectral data in `.csv` format. 
At the beginning of each classification process, an entry is created in the 
`smp` directory named with the current timestamp as well as the `MODEL` type which 
is going to be used. 

```{r classification-header, eval = FALSE}
MODEL = "BASE"
TYPE = "FUSION"
FORMAT = ".txt"

TIME = format(Sys.time(),"%Y%m%d_%H%M")
root = paste0(smp,TIME,"_",TYPE)
plots = paste0(root,"/plots")
raw = paste0(root,"/files")
dir.create(root)
dir.create(plots)
dir.create(raw)
model = paste0(mod,MODEL)

```

## Reading Data

The database and the sample data is read into R. The reflectance values
of the sample data are resampled to the spectral resolution of the database and then
a baseline correction is applied, using the procedure of @Primpke2018.
Finally, all machine-learning models are loaded from the `mod` directory.

```{r classification-data-models, eval=FALSE}

classes = readLines(paste0(ref,"classes.txt"))
data = lapply(classes,function(x){
  print(x)
  specs = read.csv(list.files(ref,full.names=T)[grep(paste("_",x,".csv",sep=""),list.files(ref))],header=T)
  return(specs)
})
data = do.call("rbind",data)
wavenumbers = readRDS(paste0(model,"/wavenumbers.rds"))

wvn = as.numeric(str_remove(names(data)[-ncol(data)],"wvn"))
index = which(wvn %in% wavenumbers)
data = data[,c(index,ncol(data))]

sampleList = list.files(smp,pattern=FORMAT,full.names = TRUE)
if (length(sampleList)==0){
  cat("No samples present in sample directory")
  #quit(status = 1)
}
Nsamples = length(sampleList)
prepSMP = function(x,wvn){
  tmp = read.table(x)
  names(tmp) = c("wavenumbers","reflectance")
  tmp = prospectr::resample2(tmp$reflectance,tmp$wavenumbers,wvn)
  return(tmp)
}

samples = lapply(sampleList,prepSMP,wavenumbers)
samples = as.data.frame(do.call("rbind",samples))
names(samples) = names(data)[-ncol(data)]

dummy = as.matrix(samples)
baselineDummy = baseline(dummy,method="rfbaseline",span=NULL,NoXP=64,maxit=c(10))
#baselineDummy = baseline(dummy, method="rollingBall", wm = 500, ws = 500 )
spectra = getCorrected(baselineDummy)
samples = as.data.frame(spectra)


files = list.files(model, full.names = TRUE)
rfModRaw = readRDS(files[grep("rfModRaw.rds", files)])
rfModSG = readRDS(files[grep("rfModSG.rds", files)])
pcaRaw = readRDS(files[grep("rfModRawPCA.rds", files)])
pcaSG = readRDS(files[grep("rfModSGPCA.rds", files)])
cnnD2 = keras::load_model_hdf5(files[grep("cnnD2",files)])
cnnND2 = keras::load_model_hdf5(files[grep("cnnND2", files)])

```

## Classification

For the decision fusion, the samples are pre-processed according to the expected input
for the different models (see [here](https://github.com/goergen95/polymeRID/blob/master/classification.R))
and the wavenumbers in the C02 window (2200 to 2420 1/cm) are set to 0. Each model 
is then used to predict an output for the samples. The decision fusion
takes places by combining the probability outputs from each model. In addition,
all non-synthetic polymer classes are merged to a broader class named `OTHER`.

```{r classification-prediction, eval= FALSE}
 # predicting
  classRFRaw = as.character(stats::predict(rfModRaw, pcaRAW))
  propRFRaw =  stats::predict(rfModRaw, pcaRAW, type = "prob")
  classRFSG = as.character(stats::predict(rfModSG, pcaSG))
  propRFSG = stats::predict(rfModSG, pcaSG, type = "prob")
  classCNND2 = as.character(classes[keras::predict_classes(cnnD2, x_sampleD2)+1])
  propCNND2 = keras::predict_proba(cnnD2, x_sampleD2)
  classCNNND2 = as.character(classes[keras::predict_classes(cnnND2, x_sampleND2)+1])
  propCNNND2 = keras::predict_proba(cnnND2, x_sampleND2)

  # restructuring results
  probs = (propRFRaw + propRFSG + propCNND2 + propCNNND2) / 4
  pred = lapply(1:nrow(probs), function(x){
    which.max(probs[x,])
  })
  predVals = lapply(1:nrow(probs), function(x){
    probs[x,unlist(pred)[x]]
  })
  hits = lapply(1:nrow(probs), function(x){
    hits = sort(probs[x, ], decreasing = T)[1:3]
  })

  predVals = unlist(predVals)
  pred = names(unlist(pred))
  pred[which(pred %in% c("FIBRE","FUR","WOOD"))] = "OTHER"
  results = data.frame(id = ids, class = pred, prob = predVals, level = rep(0,Nsamples))
```

## Generating Output

In conclusion, several outputs are created for the user to assess the classification results.
Individual plots with the three classes showing the highest probability
are created for each sample. Furthermore, a data 
frame named `results` containing information on the level of agreement for 
the class with the highest probability is written to disk to allow a quick 
assessment of the classification process. The level of agreement is based on the
fused classification probability, labeling probabilities below 0.5 as "no agreement" 
and increasing the agreement level every 10% up to >0.90 labeled as "very high agreement".
The plots are saved to a `plot` directory in the current classification directory
and can be used to manually assess the classification.

```{r classification-plots, eval=FALSE}
ids = list.files(smp,pattern = FORMAT)
ids = str_remove(ids, FORMAT)

 for (id in 1:length(ids)){
    hit = hits[[id]]
    classes = names(hit)
    values = as.numeric(hit)
    sample = as.data.frame(t(samples[id,]))
    sample$wavenumbers = wavenumbers
    sample[which(wavenumbers<=2420 & wavenumbers>=2200),] = 0
    names(sample) = c( "reflectance", "wavenumbers")
    if(values[1] < .5) level = "no agreement"
    if(values[1] >= .5 & values[1] < .6) level = "very low agreement"
    if(values[1] >= .6 & values[1] < .7) level = "low agreement"
    if(values[1] >= .7 & values[1] < .8) level = "medium agreement"
    if(values[1] >= .8 & values[1] < .9) level = "high agreement"
    if(values[1] >= .9) level = "very high agreement"
    results$level[id] = level

    annotation = paste0(level,"\n",
                        classes[1], ": ", round(values[1], 3), "\n",
                        classes[2], ": ", round(values[2], 3), "\n",
                        classes[3], ": ", round(values[3], 3))
    class1 = samplePlot(data = data, sample = sample, class = classes[1], prob = annotation, name = ids[id])
    class2 = samplePlot(data = data, sample = sample, class = classes[2])
    class3 = samplePlot(data = data, sample = sample, class = classes[3])
    multiclass = gridExtra::grid.arrange(class1,class2,class3)
    ggsave(plot=multiclass,file=paste0(plots,"/",ids[id],"_probClasses.png"),dpi=300,device="png",units="cm",width=50,height=30)
  }
  write.csv(results, paste0(root, "/results_",TIME,".csv"))
```


## Results

Here, we evaluate the decision fusion models by some environmental samples which 
were provided by Sarah Brüning and Frauke von den Driesch. 
The overall classification results, as they are written to
the `results.csv`, are presented below.

```{r classification-classification, warning=FALSE, message=FALSE, echo=FALSE}
results = read.csv(paste0(smp,"20190819_1910_FUSION/results_20190819_1910.csv"))
knitr::kable(results, caption = "**Tab. 1**: Classification results for 14 sample spectra.")
```

Additionally, two exemplary plots overlaying the spectra of two
samples with mean spectra of the classified polymer in the database are shown. 
Note that the first plot corresponds to row 5 in the table above, the second plot to line 13. 

```{r classification-sample-plot, warning=FALSE, message=FALSE, echo=FALSE}
MODEL = "BASE"
TYPE = "FUSION"
FORMAT = ".txt"
library(plotly)
wavenumbers = readRDS(paste0(mod,"BASE/wavenumbers.rds"))
sampleList = list.files(smp,pattern=FORMAT,full.names = TRUE)
if (length(sampleList)==0){
  cat("No samples present in sample directory")
  #quit(status = 1)
}
Nsamples = length(sampleList)
prepSMP = function(x,wvn){
  tmp = read.table(x)
  names(tmp) = c("wavenumbers","reflectance")
  tmp = prospectr::resample2(tmp$reflectance,tmp$wavenumbers,wvn)
  return(tmp)
}

samples = lapply(sampleList,prepSMP,wavenumbers)
samples = as.data.frame(do.call("rbind",samples))
names(samples) = names(data)[-ncol(data)]

dummy = as.matrix(samples)
baselineDummy = baseline(dummy,method="rfbaseline",span=NULL,NoXP=64,maxit=c(10))
#baselineDummy = baseline(dummy, method="rollingBall", wm = 500, ws = 500 )
spectra = getCorrected(baselineDummy)
samples = as.data.frame(spectra)
samples[,which(wavenumbers<=2420 & wavenumbers>=2200)] = 0

sample = read.table(sampleList[5])
sample[which(wavenumbers<=2420 & wavenumbers>=2200),] = 0
names(sample) = c("wavenumbers", "reflectance")


p = samplePlot(data, sample, class = "HDPE")
p = ggplotly(p)

sample = read.table(sampleList[13])
sample[which(wavenumbers<=2420 & wavenumbers>=2200),] = 0
names(sample) = c("wavenumbers", "reflectance")
p2 = samplePlot(data, sample, class = "PA")
p2 = ggplotly(p2)
p
p2

```



