---
title: "Classification"
author: Darius Goergen
site: workflowr::wflow_site
output:
  workflowr::wflow_html:
    toc: true
editor_options:
  chunk_output_type: console
bibliography: library.bib
link-citations: yes
---

```{r calibration-setup, echo=FALSE, warning=FALSE, message=FALSE}
source("code/setup_website.R")
source("code/functions.R")
data = read.csv(file = paste0(ref, "reference_database.csv"), header = TRUE)
```

## Overview

The overall aim of this project was to ease the process of classifying the spectra
of particles found in environmental samples. To this end, we created a script which
takes `.txt` files in the `smp` directory and classifies them based on the decision
fusion explained [here](calibration.html).
The script expects on file for each sample. The names of the files need to be unique
and will be used as an identifier in the output. Also, it is expected that the file
consists of two columns. The first columns indicates the wavenumbers of the sample
as an numeric, while the second column holds information about the reflectance value.

## Preparing Classification

The head of the classification script contains some variables that might be changed.
This is important if extensions were made to the data base and a new calibration was done.
In this case, the directory containing the models can be specified with the `MODEL`
variable. For now, however, the variables hints towards the `BASE` models which were created
during the first calibration. Also, the type of the classification can be specified in the
`TYPE` variable. Currently, the fusion of all four models is chosen, but any single model
could be selected. Finally, the `FORMAT` variables indicates the file extension of the sample
data, since it would be also possible to provide spectral data in `.csv` format. 
At the beginning of each classification process, a directory is created in the 
`smp` directory which contains the current date and time as well as the `MODEL` type which 
was used. 

```{r classification-header, eval = FALSE}
MODEL = "BASE"
TYPE = "FUSION"
FORMAT = ".txt"

TIME = format(Sys.time(),"%Y%m%d_%H%M")
root = paste0(smp,TIME,"_",TYPE)
plots = paste0(root,"/plots")
raw = paste0(root,"/files")
dir.create(root)
dir.create(plots)
dir.create(raw)
model = paste0(mod,MODEL)

```

## Reading Data

Then, the data base and the sample data is read into R. The reflectance values
of the sample data are resampled to the spectral resolution of the data base and then
a baseline correction is applied, following the procedure of @Primpke2018.
Finally, all machine learning models are loaded.

```{r classification-data-models, eval=FALSE}

classes = readLines(paste0(ref,"classes.txt"))
data = lapply(classes,function(x){
  print(x)
  specs = read.csv(list.files(ref,full.names=T)[grep(paste("_",x,".csv",sep=""),list.files(ref))],header=T)
  return(specs)
})
data = do.call("rbind",data)
wavenumbers = readRDS(paste0(model,"/wavenumbers.rds"))

wvn = as.numeric(str_remove(names(data)[-ncol(data)],"wvn"))
index = which(wvn %in% wavenumbers)
data = data[,c(index,ncol(data))]

sampleList = list.files(smp,pattern=FORMAT,full.names = TRUE)
if (length(sampleList)==0){
  cat("No samples present in sample directory")
  #quit(status = 1)
}
Nsamples = length(sampleList)
prepSMP = function(x,wvn){
  tmp = read.table(x)
  names(tmp) = c("wavenumbers","reflectance")
  tmp = prospectr::resample2(tmp$reflectance,tmp$wavenumbers,wvn)
  return(tmp)
}

samples = lapply(sampleList,prepSMP,wavenumbers)
samples = as.data.frame(do.call("rbind",samples))
names(samples) = names(data)[-ncol(data)]

dummy = as.matrix(samples)
baselineDummy = baseline(dummy,method="rfbaseline",span=NULL,NoXP=64,maxit=c(10))
#baselineDummy = baseline(dummy, method="rollingBall", wm = 500, ws = 500 )
spectra = getCorrected(baselineDummy)
samples = as.data.frame(spectra)


files = list.files(model, full.names = TRUE)
rfModRaw = readRDS(files[grep("rfModRaw.rds", files)])
rfModSG = readRDS(files[grep("rfModSG.rds", files)])
pcaRaw = readRDS(files[grep("rfModRawPCA.rds", files)])
pcaSG = readRDS(files[grep("rfModSGPCA.rds", files)])
cnnD2 = keras::load_model_hdf5(files[grep("cnnD2",files)])
cnnND2 = keras::load_model_hdf5(files[grep("cnnND2", files)])

```

## Classification

For the decision fusion, the samples are preprocessed according to the expected input
for the different models (see [here](https://github.com/goergen95/polymeRID/blob/master/classification.R))
and the wavenumbers in the C02 window (2420 to 2200 1/cm) are set to 0. Then, 
each model is used to predict an output for the samples and the decision fusion
takes places by combining the probability outputs from all models. Additionally,
all non-synthetic polymer classes are merged to a broader class called `OTHER`.

```{r classification-prediction, eval= FALSE}
 # predicting
  classRFRaw = as.character(stats::predict(rfModRaw, pcaRAW))
  propRFRaw =  stats::predict(rfModRaw, pcaRAW, type = "prob")
  classRFSG = as.character(stats::predict(rfModSG, pcaSG))
  propRFSG = stats::predict(rfModSG, pcaSG, type = "prob")
  classCNND2 = as.character(classes[keras::predict_classes(cnnD2, x_sampleD2)+1])
  propCNND2 = keras::predict_proba(cnnD2, x_sampleD2)
  classCNNND2 = as.character(classes[keras::predict_classes(cnnND2, x_sampleND2)+1])
  propCNNND2 = keras::predict_proba(cnnND2, x_sampleND2)

  # restructuring results
  probs = (propRFRaw + propRFSG + propCNND2 + propCNNND2) / 4
  pred = lapply(1:nrow(probs), function(x){
    which.max(probs[x,])
  })
  predVals = lapply(1:nrow(probs), function(x){
    probs[x,unlist(pred)[x]]
  })
  hits = lapply(1:nrow(probs), function(x){
    hits = sort(probs[x, ], decreasing = T)[1:3]
  })

  predVals = unlist(predVals)
  pred = names(unlist(pred))
  pred[which(pred %in% c("FIBRE","FUR","WOOD"))] = "OTHER"
  results = data.frame(id = ids, class = pred, prob = predVals, level = rep(0,Nsamples))
```

## Generating Output

Finally, an individual plot with the three classes showing the highest probability
is created for each sample. Additionally, a data frame called `results` containing
information on the level of agreement for the class with the highest probability
is written to disk to allow a quick assessment of the classification process.
The level of agreement is based on the fused classification probability, labeling
probabilities below 0.5 as "now agreement" and increasing the agreement level every
10% up to >0.90 labeled as "very high agreement".
The plots are saved to a `plot` directory in the current classification directory
and can be used to manually assess the classification.

```{r classification-plots, eval=FALSE}
ids = list.files(smp,pattern = FORMAT)
ids = str_remove(ids, FORMAT)

 for (id in 1:length(ids)){
    hit = hits[[id]]
    classes = names(hit)
    values = as.numeric(hit)
    sample = as.data.frame(t(samples[id,]))
    sample$wavenumbers = wavenumbers
    sample[which(wavenumbers<=2420 & wavenumbers>=2200),] = 0
    names(sample) = c( "reflectance", "wavenumbers")
    if(values[1] < .5) level = "no agreement"
    if(values[1] >= .5 & values[1] < .6) level = "very low agreement"
    if(values[1] >= .6 & values[1] < .7) level = "low agreement"
    if(values[1] >= .7 & values[1] < .8) level = "medium agreement"
    if(values[1] >= .8 & values[1] < .9) level = "high agreement"
    if(values[1] >= .9) level = "very high agreement"
    results$level[id] = level

    annotation = paste0(level,"\n",
                        classes[1], ": ", round(values[1], 3), "\n",
                        classes[2], ": ", round(values[2], 3), "\n",
                        classes[3], ": ", round(values[3], 3))
    class1 = samplePlot(data = data, sample = sample, class = classes[1], prob = annotation, name = ids[id])
    class2 = samplePlot(data = data, sample = sample, class = classes[2])
    class3 = samplePlot(data = data, sample = sample, class = classes[3])
    multiclass = gridExtra::grid.arrange(class1,class2,class3)
    ggsave(plot=multiclass,file=paste0(plots,"/",ids[id],"_probClasses.png"),dpi=300,device="png",units="cm",width=50,height=30)
  }
  write.csv(results, paste0(root, "/results_",TIME,".csv"))
```


## Results

Here, we evaluate the decision fusion models by some environmental samples which 
were provided by Sarah Brüning and Frauke von den Driesch. 
First we can take a look at the general classification results as it is written to
the `results.csv`.
```{r classification-classification, warning=FALSE, message=FALSE, echo=FALSE}
results = read.csv(paste0(smp,"20190819_1910_FUSION/results_20190819_1910.csv"))
knitr::kable(results, caption = "**Tab. 1**: Classification results for 14 sample spectra.")
```

Additionally, we included two exemplary plots, overlaying the spectrum of two
samples with the mean spectrum of the prediction. Note that the firs plot corresponds
to row 5 in the table above, the second plot to line 13. 

```{r classification-sample-plot, warning=FALSE, message=FALSE, echo=FALSE}
MODEL = "BASE"
TYPE = "FUSION"
FORMAT = ".txt"
library(plotly)
wavenumbers = readRDS(paste0(mod,"BASE/wavenumbers.rds"))
sampleList = list.files(smp,pattern=FORMAT,full.names = TRUE)
if (length(sampleList)==0){
  cat("No samples present in sample directory")
  #quit(status = 1)
}
Nsamples = length(sampleList)
prepSMP = function(x,wvn){
  tmp = read.table(x)
  names(tmp) = c("wavenumbers","reflectance")
  tmp = prospectr::resample2(tmp$reflectance,tmp$wavenumbers,wvn)
  return(tmp)
}

samples = lapply(sampleList,prepSMP,wavenumbers)
samples = as.data.frame(do.call("rbind",samples))
names(samples) = names(data)[-ncol(data)]

dummy = as.matrix(samples)
baselineDummy = baseline(dummy,method="rfbaseline",span=NULL,NoXP=64,maxit=c(10))
#baselineDummy = baseline(dummy, method="rollingBall", wm = 500, ws = 500 )
spectra = getCorrected(baselineDummy)
samples = as.data.frame(spectra)
samples[,which(wavenumbers<=2420 & wavenumbers>=2200)] = 0

sample = read.table(sampleList[5])
sample[which(wavenumbers<=2420 & wavenumbers>=2200),] = 0
names(sample) = c("wavenumbers", "reflectance")


p = samplePlot(data, sample, class = "HDPE")
p = ggplotly(p)

sample = read.table(sampleList[13])
sample[which(wavenumbers<=2420 & wavenumbers>=2200),] = 0
names(sample) = c("wavenumbers", "reflectance")
p2 = samplePlot(data, sample, class = "PA")
p2 = ggplotly(p2)
p
p2

```



