---
title: "Preparation"
author: Darius Goergen
site: workflowr::wflow_site
output:
  workflowr::wflow_html:
    toc: false
editor_options:
  chunk_output_type: console
bibliography: library.bib
link-citations: yes
---

```{r setup, include=FALSE, warning=FALSE, message=FALSE}
source("code/setup_website.R")
source("code/functions.R")
```

For the current project we used a data base published by @Primpke2018 online. 
The data base can be downloaded 
[here](https://static-content.springer.com/esm/art%3A10.1007%2Fs00216-018-1156-x/MediaObjects/216_2018_1156_MOESM2_ESM.xlsx).
The authors state the samples were collected based on the FTIR-spectrometer Bruker Tensor 27 System
for the spectral range 4000 to 40 1/cm. Additionally, some data of polymer-based
fibers and spectra of biological origins were received from the [Bremer Faserinstitut](https://www.faserinstitut.de/).
During preprocessing, they applied a concave rubberband correction based on 10 
iterations and 64 baseline points. They also excluded the C02 band between
2420 to 2200 1/cm by setting the data points to 0. This should be kept in mind,
since the inclusion of additional reference samples requires the same procedure
for the data base to stay in a consistent state. The data provided by @Primpke2018 
shows a spectral resolution of 2.1 1/cm. Additional reference samples need be
resampled to the same spectral resolution.

To ensure this, the data base was read into R and the wave number index was saved 
in a separate file for future use.
```{r read-in-data}
library(openxlsx)
url = "https://static-content.springer.com/esm/art%3A10.1007%2Fs00216-018-1156-x/MediaObjects/216_2018_1156_MOESM2_ESM.xlsx"

data = openxlsx::read.xlsx(url)
# extract wavenumbers from first row
wavenumbers = as.numeric(names(data)[2:1864])
# saving wavenumbers to reference sample directory
saveRDS(wavenumbers, paste0(ref, "wavenumbers.rds"))
```

We can take a look at the distribution of the different classes found in the data base.
Here, I only print the 20 most common classes since there are a lot of reference samples
only found once.
```{r analyse-classes}
data$Abbreviation = as.factor(data$Abbreviation)
summary(data$Abbreviation)[1:10]

```

Because we are interested in assigning the right class to potential plastic particles, 
the most important classes found in the data base 
to us are the ones of artificial origin. However, sometimes also particles of 
biological origins are subject to spectral analysis, because they resemble the 
appearance of micro plastics. Any machine learning algorithm trained only with
reference samples from plastics would eventually assign a plastic class also 
to the particles of biological origin. To reduce the error of false positives, 
we include some of the samples of biological origin based on broad classes.

```{r retrieve-target-classes}
# furs and wools
indexFur = grep("fur", data$Abbreviation)
indexWool = grep("wool", data$Abbreviation)
furs = data[c(indexFur, indexWool), ]
furs = furs[ , c(2:1864)] # leave out index column
names(furs) = paste("wvn", wavenumbers, sep="")
furs$class = "FUR"

# fibres
indexFibre = grep("fibre", data$Abbreviation)
fibre = data[indexFibre, ]
fibre = fibre[ , c(2:1864)] # leave out index column
names(fibre) = paste("wvn", wavenumbers, sep="")
fibre$class = "FIBRE"

# wood
indexWood = grep("wood", data$Abbreviation)
wood = data[indexWood, ]
wood = wood[ , c(2:1864)] # leave out index colums
names(wood) = paste("wvn", wavenumbers, sep="")
wood$class = "WOOD"

# synthetic polymers
polyIndex = which(data$`Natural./Synthetic` =="synthetic polymer")
syntPolymer = data[polyIndex,]
counts = summary(syntPolymer$Abbreviation)
polyNames = names(counts)[1:10] # only major polymers
syntPolymer = syntPolymer[which(syntPolymer$Abbreviation %in%  polyNames) , ]
classes = droplevels(syntPolymer$Abbreviation)
syntPolymer = syntPolymer[ , c(2:1864)] # leave out index column
names(syntPolymer) = paste("wvn",wavenumbers,sep="")
syntPolymer$class = as.character(classes)

# lets group together some synthetic polymer classes
syntPolymer$class[grep("Nylon",syntPolymer$class)] = "PA"
```

We can now bind the reference samples together and take a look at the distribution
of classes in the resulting data frame.

```{r binding-data}
data = rbind(furs,wood,fibre,syntPolymer) 
data$class = as.factor(data$class)
summary(data$class)

```
We see that we have in total 93 (53%) reference samples of plastic polymers and 
44 (47%) of biological origin. Within the plastic samples, we find that the data 
is very balanced with no single class showing less than 7 samples. For the samples
of biological origin, however, the class fiber dominates the class distribution.
This could prove as an disadvantage if a machine learning algorithm picks up 
this unbalance by minimizing its error rate simply by more frequently predicting
the fiber class. At this point, we will leave the resulting data base as it is 
and save it to disk.
We save the data in individual files as well as in a comprehensive data base in 
csv format. This way we ensure that later extensions to the data base  easily to
manage.

```{r write-data}
write.csv(data, file = paste0(ref, "reference_database.csv"), row.names=FALSE)

# writing class control file
classIndex = as.character(unique(data$class))

for (class in classIndex){
  tmp = data[data$class==class , ]
  write.csv(tmp, file = paste0(ref, "reference_", class, ".csv"), row.names=FALSE)
}

write(classIndex, paste0(ref, "classes.txt"))
```


## Literature used on this page

