---
title: "Preparation"
author: Darius Goergen
site: workflowr::wflow_site
output:
  workflowr::wflow_html:
    toc: true
editor_options:
  chunk_output_type: console
bibliography: library.bib
link-citations: yes
runtime: shiny
---

```{r setup, include=FALSE, warning=FALSE, message=FALSE}
source("code/setup_website.R")
source("code/functions.R")
```

## Reference Data
For this project we used a data base published by @Primpke2018 online. 
The data base can be downloaded 
[here](https://static-content.springer.com/esm/art%3A10.1007%2Fs00216-018-1156-x/MediaObjects/216_2018_1156_MOESM2_ESM.xlsx).
The authors state that the samples were collected based on the FTIR-spectrometer Bruker Tensor 27 System
for the spectral range 4000 to 40 1/cm. Additionally, some data of polymer-based
fibers and spectra of biological origins were received from the [Bremer Faserinstitut](https://www.faserinstitut.de/).
During pre-processing, they applied a concave rubberband correction based on ten 
iterations and 64 baseline points. They also excluded the C02 band between
2200 to 2420 1/cm by setting the data points to 0. This should be kept in mind,
since the inclusion of additional reference samples requires the same procedure
for the database to stay in a consistent state. The data shows a spectral resolution 
of 2.1 1/cm. Additional reference samples need to be resampled to the same spectral 
resolution and the same baseline correction should be applied.

To ensure consistency, the data base was read into `R` and the wavenumbers were saved 
in a separate file for the future use of resampling additional reference spectra.

```{r read-in-data}
library(openxlsx)
url = "https://static-content.springer.com/esm/art%3A10.1007%2Fs00216-018-1156-x/MediaObjects/216_2018_1156_MOESM2_ESM.xlsx"

data = openxlsx::read.xlsx(url)
# extract wavenumbers from first row
wavenumbers = as.numeric(names(data)[2:1864])
# saving wavenumbers to reference sample directory
saveRDS(wavenumbers, paste0(ref, "wavenumbers.rds"))
```

An important feature of any data base is the distribution of different classes.
Here, we only print the 20 most common classes, because there are a lot of reference samples
only found once or twice within the database. 

```{r analyse-classes}
data$Abbreviation = as.factor(data$Abbreviation)
summary(data$Abbreviation)[1:10]

```

## Construction of the Database
We are interested in assigning the correct class to potential plastic particles. 
The most important classes to us found in the database are the ones of 
artificial polymer origin. However, sometimes particles of biological origins
will also be subject to a spectral analysis, because they resemble the appearance of 
microplastic in environmental samples. Any machine-learning algorithm trained 
only with reference samples from plastics would eventually assign one of these classes
to the particles of biological origins. It will only assign the class with the
greatest similarity to the classes it has learned. This can lead to so-called false 
positive errors. To reduce the occurence of false positives we include some of the 
samples of biological origin as well. We summarize these samples to broader classes.

```{r retrieve-target-classes}
# furs and wools
indexFur = grep("fur", data$Abbreviation)
indexWool = grep("wool", data$Abbreviation)
furs = data[c(indexFur, indexWool), ]
furs = furs[ , c(2:1864)] # leave out index column
names(furs) = paste("wvn", wavenumbers, sep="")
furs$class = "FUR"

# fibres
indexFibre = grep("fibre", data$Abbreviation)
fibre = data[indexFibre, ]
fibre = fibre[ , c(2:1864)] # leave out index column
names(fibre) = paste("wvn", wavenumbers, sep="")
fibre$class = "FIBRE"

# wood
indexWood = grep("wood", data$Abbreviation)
wood = data[indexWood, ]
wood = wood[ , c(2:1864)] # leave out index colums
names(wood) = paste("wvn", wavenumbers, sep="")
wood$class = "WOOD"

# synthetic polymers
polyIndex = which(data$`Natural./Synthetic` =="synthetic polymer")
syntPolymer = data[polyIndex,]
counts = summary(syntPolymer$Abbreviation)
polyNames = names(counts)[1:10] # only major polymers
syntPolymer = syntPolymer[which(syntPolymer$Abbreviation %in%  polyNames) , ]
classes = droplevels(syntPolymer$Abbreviation)
syntPolymer = syntPolymer[ , c(2:1864)] # leave out index column
names(syntPolymer) = paste("wvn",wavenumbers,sep="")
syntPolymer$class = as.character(classes)

# lets group together some synthetic polymer classes
syntPolymer$class[grep("Nylon",syntPolymer$class)] = "PA"
```

## Class Distribution

We now bind the reference samples together and take a look at the distribution
of classes in the resulting data frame, which is the concrete database used for the 
following calculations.

```{r binding-data}
data = rbind(furs,wood,fibre,syntPolymer) 
data$class = as.factor(data$class)
summary(data$class)

```

In total, 93 (53%) reference samples of plastic polymers are present in the database
and 44 (47%) of biological origin. Within the plastic samples, we found that the data 
is very balanced with no single class showing less than seven samples. For the samples
of biological origin, however, the class `FIBRE` dominates the distribution.
This could prove a disadvantage if a machine-learning algorithm picks up 
this unbalance by minimizing its error-rate simply by more frequently predicting
the `FIBRE` class. At this point, we will leave the resulting data-base as it is 
and save it to disk.
We save the data in individual files as well as, in a comprehensive file in 
`.csv` format. This way we ensure that later extensions to the database are more 
easily manageable.

```{r write-data}
write.csv(data, file = paste0(ref, "reference_database.csv"), row.names=FALSE)

# writing class control file
classIndex = as.character(unique(data$class))

for (class in classIndex){
  tmp = data[data$class==class , ]
  write.csv(tmp, file = paste0(ref, "reference_", class, ".csv"), row.names=FALSE)
}

write(classIndex, paste0(ref, "classes.txt"))
```

## Spectra of Reference Samples
One can find below the spectra found within the database. By selecting an entry
in the drop-down menu the plot of the respective class will be shown. The solid
grey line in the center of the plot indicates the mean value for all samples of 
the respective wavenumbers. The grey ribbon indicates the standard deviation from
that mean value, while the dashed lines show the minimum and the maximum values,
respectively. 

```{r preparation-shiny-app, echo=FALSE, warning=FALSE, message=FALSE}
knitr::include_app("https://goergen95.shinyapps.io/shiny_apps/", height = "500px")
```


## Citations on this page

