---
title: "Support Vector Machine"
author: Darius Goergen
site: workflowr::wflow_site
output:
  workflowr::wflow_html:
    toc: false
editor_options:
  chunk_output_type: console
bibliography: library.bib
link-citations: yes
---

```{r setup, include=FALSE, warning=FALSE, message=FALSE}
source("code/setup_website.R")
source("code/functions.R")
```

Support Vector Machine (SVM) is a non-parametric classification method which is
initially designed for binary classification problems and which was developed in 
its current form by @Boser:1992:TAO:130385.130401. A detailed overview of the 
SVM algorithm is found in @Burges1998. The principal idea behind SVM is to find 
an optimal hyperplane which separates two classes from another by the largest
possible margin. The alogrithm is optimised by iterativly maximizing this margin 
but while only considering the closest observations of both classes to the margin. 
These specific observations are also called support-vectors. Multidimensional 
data can be processed by mapping the data into a higher dimensional feature 
space through a specified mapping function. This function is called kernel 
function, and mainly four different groups are used: linear, polynomial, radial 
and sigmoid functions [@Burges1998]. In this project only the radial basis 
function was used. Multi-class problems are addressed by calculating an optimal 
margin following the one class-against-all pattern and conducting a majority vote 
and the end of the calculations. SVM need some tuning parameters. These are the 
regularization parameter `C` and the kernel width `γ`. The regularization 
parameter is also called penalty value, as it is a constant giving penalty to 
misclassified observations. There optimal values might change if different 
representations of the data are presented to the algorithm.

Different levels of data reprocessing might emphasize different features of 
the patterns to be learned by an algorithm. To grasp this, different transformations
of the data were presented to the SVM algorithm. Additionally, the raw data signal
was jittered to test which transformation might prove beneficial in delivering 
high classification accuracies even in the presence of noise. To test this, 
we define a function which adds noise to the raw data and returns a list with
the number of elements equal to the levels of noise applied.


```{r svm-addNoise}
addNoise = function(data, levels = c(0), category="class"){
  data.return = list()
  index = which(names(data) == category)
  for (n in levels){
    tmp = as.matrix(data[ , -index])
    if (n == 0){
      tmp = data
    }else{
      tmp = as.data.frame(jitter(tmp, n))
      tmp[category] = data[category]
    }
    data.return[[paste("noise", n, sep="")]] = tmp
  }
  return(data.return)
}

data = read.csv(file = paste0(ref, "reference_database.csv"), header = TRUE)
noisy_data = addNoise(data, levels = c(0,10,100,250,500), category = "class")

# indivitual elements can be selected by using [[ and refering to the index or the name
head(noisy_data[["noise100"]])[1:3,1:3]
```

Then, in another user defined function which uses the `noisy_data` objected as input
specified data transformations are applied. These are normalization which centers
and scales the input data, as well as different forms of the Savitkiy-Golay filter [@Savitzky1964]
and first and second derivative of a raw spectrum. The functions iterates through
the noise level elements in the `noisy_data` object and returns each specified 
transformation in a list element below the noise level. The exemplary function below
applies the preprocessing for normalization, standard filtering and first 
derivative only. The implementation of the function used in the project
can be found [here](https://github.com/goergen95/polymeRID/blob/master/code/functions.R#12).

```{r svm-TrainigSet}
createTrainingSet = function(data, category = "class",
                             SGpara = list(p=3,w=11), lag = 15){
  
  data.return = list()
  for (noise in names(data)){
    
    tmp = as.data.frame(data[[noise]])
    classes = tmp[,category]
    tmp = tmp[!names(tmp) %in% category]
    
    # original data
    data.return[[noise]][["raw"]] = as.data.frame(data[[noise]])
    
    # normalised data
    data_norm = preprocess(tmp, type="norm")
    data_norm[category] = classes
    data.return[[noise]][["norm"]] = data_norm
    
    # SG-filtered data
    data_sg = preprocess(tmp, type="sg", SGpara = SGpara)
    data_sg[category] = classes
    data.return[[noise]][["sg"]] = data_sg
    
    # first derivative of original data
    data_rawd1 = preprocess(tmp, type="raw.d1", lag = lag)
    data_rawd1[category] = classes
    data.return[[noise]][["raw.d1"]] = data_rawd1
    
  }
  return(data.return)
}

# applying the function
test_dataset = createTrainingSet(noisy_data, category = "class")

# individual transformations at a certain noise level can be accessed with [[
head(test_dataset[["noise500"]][["raw.d1"]])[1:3,1:3]
```

The data base of [@Primpke2018] currently shows 1863 variables for each observations.
Most of these data points do not bear relevant information to distinguish between
different types of particles. To shorten the computation time, one can use 
dimensionality reduction techniques such as principal component analysis (PCA).
PCA also has been used to transform spectral data of micro-plastics in marine 
ecosystems before [@Jung2018;@Lorenzo-Navarro2018]. PCA basically takes the input
data for a given number of observation and by performing a orthogonal transformation
to the data transforms these possible correlated variables to uncorrelated 
principal components. This way, both redundancies in the data as well as possible
noise can be accounted for. PCA previously has been successfully applied to
FTIR-spectrometer data [@Hori2003;@Nieuwoudt2004;@Mueller2013;@Ami2013;Fu2014].
Simultaniously, the number of variables can be significantly reduced by applying
PCA and thus speeding up the training process. Below we will apply a PCA to the
raw data as an example.
```{r PCA-raw , message=FALSE}
library(factoextra)
tmp = test_dataset[["noise0"]][["raw"]]
pca = prcomp(tmp[ ,-1864]) # omitting class variable
var_info = factoextra::get_eigenvalue(pca)
# setting a threshold of 99% explained variance
threshold = 99
thresInd = which(var_info$cumulative.variance.percent>=threshold)[1]
pca_data = pca$x[,1:thresInd]
```
We can use the index variable `thresInd` we just defined to take a look upon
all the principal components which explain 99% of the variance in the data set.

```{r pca-table, echo=FALSE}
library(knitr)
knitr::kable(var_info[1:thresInd,])
```

We effectivly reduced the number of variables from 1683 to 15 which still bear 
99% of the variance we can find in the original data set. However, when it comes
to machine learning, it is important to realise that this new dataset is not fit
to be used in a training process. If we now randomly split the observations into
training and test, we effectivly mix up these two sets because information of 
the test set has already influenced the outcome of the PCA. Therefor, the data set
need to be split beforhand of the PCA. The analysis is done on the training data 
only and then the same orthogonal transformations will be applied to the test data.
This way it can be ensured that the test set is truely independent from the 
training process.
Here, we apply a 10-fold cross-validation which is repeated 5 times. The following 
code takes a complete data set as input, applies a splitting function from the
`caret` package and then builds the PCA upon the the test set and finally applies
the same transformation to the test set. We apply it for the raw data only.
Also, we randomly split the data to 50% training and 50% test.

```{r PCA_loop}
folds = 10
repeats = 5
split_percentage = 0.5
threshold = 99
tmp = test_dataset[["noise0"]][["raw"]]

set.seed(42) # ensure reproducibility
fold_index = lapply(1:repeats, caret::createDataPartition, y=tmp$class,
                   times = folds, p = split_percentage)
fold_index = do.call(c, fold_index)

pcaData = list()
for (rep in 1:repeats){
  rep_index = fold_index[(rep*folds-folds+1):(rep*folds)] # jumps to the correct number of folds forward in each repeat
  
  pcadata_fold = lapply(1:folds,function(x){
    
    # splitting for current fold
    training = tmp[unlist(rep_index[x]),]
    validation = tmp[-unlist(rep_index[x]),]
    
    # keep response
    responseTrain = training$class
    responseVal = validation$class
    
    # apply PCA
    pca = prcomp(training[,1:1863])
    varInfo = factoextra::get_eigenvalue(pca)
    thresInd = which(varInfo$cumulative.variance.percent >= threshold)[1]
    pca_training = pca$x[ ,1:thresInd]
    pca_validation = predict(pca, validation)[ ,1:thresInd]
    
    training = as.data.frame(pca_training)
    training$response = responseTrain
    validation = as.data.frame(pca_validation)
    validation$response = responseVal
    foldtmp = list(training, validation)
    names(foldtmp) = c("training","validation")
    return(foldtmp)
  })
  names(pcadata_fold) = paste("fold", 1:folds, sep ="")
  pcaData[[paste0("repeat",rep)]] = pcadata_fold
}

```
We now have a list object with the number of elements equivalent to the repeats.
Below each repeat element we can access the individual folds.
There we find two elements which we can access by refering to `"training"`
and `"testing"`.
```{r PCA_access}
pcaData[["repeat5"]][["fold10"]][["training"]][1:3,1:3]
pcaData[["repeat5"]][["fold10"]][["validation"]][1:3,1:3]

summary(pcaData[["repeat5"]][["fold10"]][["training"]]$response)
summary(pcaData[["repeat5"]][["fold10"]][["validation"]]$response)
```
For the SVM algorithm we also implementd a simple search pattern for optimal 
paramters for the regularization parameter `C` and the kernel width `γ`. We did
this by applying a search grid for the paramters and calculating each possible
combination. Note that due to limits in computation capacities we restriced
the search to 25 combinations only. The code below calculates a model for each
possible combinations, evaluating it capacity to correctly classify the training
data and then chooses the optimal model to evaluate the validation data. 

```{r svm-paramter-tuning, eval = TRUE}
training = pcaData[["repeat1"]][["fold1"]][["training"]]
validation =pcaData[["repeat1"]][["fold1"]][["validation"]]
x_train = training[ ,1:ncol(training)-1]
y_train = training$response
x_test = validation[ ,1:ncol(validation)-1]
y_test = validation$response


tuneGrid = expand.grid(gamma =seq(0.1,1,0.2),cost = seq(1,5,1) )
accuracy = c()
models = list()
for ( i in 1:nrow(tuneGrid)){
  model = e1071::svm(x = x_train, y = y_train,
                    kernel = "radial",
                    gamma = tuneGrid$gamma[i],
                    cost = tuneGrid$cost[i])
  pred = predict(model, x_train)
  conf = caret::confusionMatrix(pred, y_train)
  accuracy = c(accuracy, conf$overall["Kappa"])
  models[[i]] = model
}

best_model = models[[which(accuracy == max(accuracy))[1]]]
prediction = predict(best_model, x_test)
confMat = caret::confusionMatrix(prediction, y_test)
print(confMat$table)
```

This code is integrated into a function which applies this search pattern to all 
folds for all repeats and can be found [here](https://github.com/goergen95/polymeRID/blob/master/code/functions.R#321).
Finally, we can apply this function to the different levels of preprocssing which
were discussed before and obtaining the accuracies in the `results` object.

```{r CV-apply, eval = FALSE}
source("code/functions.R")
wavenumbers = readRDS(paste0(ref,"wavenumbers.rds"))
# add noise to data
noisyData = addNoise(data,levels = c(0,10,100,250,500), category = "class")

# preprocessing
testDataset = createTrainingSet(noisyData, category = "class",
                                SGpara = list(p=3, w=11), lag=15,
                                type = c("raw", "norm", "sg", "sg.d1", "sg.d2",
                                  "sg.norm", "sg.norm.d1", "sg.norm.d2",
                                  "raw.d1", "raw.d2", "norm.d1", "norm.d2"))


types = names(testDataset[[1]])

levels = lapply(names(testDataset), function(x){
  rep(x, length(types))
})
levels = unlist(levels)

results = data.frame(level=levels, type = types, kappa = rep(0,length(levels)))

for (level in unique(levels)){
  for (type in types){

    print(paste0("Level: ",level," Type: ",type))
    tmpData = testDataset[[level]][[type]]
    tmpData[which(wavenumbers<=2420 & wavenumbers>=2200)] = 0 # setting C02 window to 0
    tmpModel = pcaCV(tmpData, folds = 10, repeats = 5, threshold = 99, metric = "Kappa", p=0.5, method="svm")
    saveRDS(tmpModel,file = paste0(output,"svm/model_",level,"_",type,"_",round(tmpModel[[1]],2),".rds"))
    results[which(results$level==level & results$type==type),"kappa"] = as.numeric(tmpModel[[1]])
    print(results)

  }
}
saveRDS(results, paste0(output,"svm/exploration.rds"))

```
```{r read-exploration, include = FALSE}
results = readRDS(paste0(output,"svm/exploration.rds"))
```

We can now take a look at the kappa scores the algorithm achived during training
for different representations of the data and at increasing noise levels.

```{r plot-results, echo = FALSE, warning=FALSE, message=FALSE}
library(plotly)
results$level = as.character(results$level)
# let's visualize the accuracy results with noise levels on x-axis and kappa score on y-axis
test = ggplot(data=results)+
  geom_line(aes(y=kappa,group=type,x=level,color=type),size=1.5)+
  ylab("Kappa score")+
  xlab("Noise level")+
  scale_color_discrete(name="Type of\nPre-Processing")+
  theme_minimal()

t = ggplotly(test)
t
```
We can observe that with higher noise ratios the kappa score is reduced significantly.
All data transformations yield to very low accuracies when the noise is increases.
In the absence of significant noise, however, the simple Savitzkiy-Golay filter,
the raw data as well as the first and second order derivatives yield to a kappa
score of about 0.75. We can look at this more mathmatically by calculating the average 
slopes of the data tranformations methods and order the data frame from low to
high slopes. Note that we only take the kappa score at noise level 0 and 10
to calculate the average slope.

```{r slope-calculation}
noise0 = results[results$level == "noise0", ]
noise10 = results[results$level == "noise10", ]
slopes = noise10$kappa - noise0$kappa 
types = unique(results$type)
df = data.frame(type = types, slope = slopes, row.names = NULL)
df = df[order(-slopes),]
```
```{r print-slope-table, echo = FALSE}
knitr::kable(df)
```

We can now confirm, that the average decrease in kappa score is the lowest for
the Savitzkiy-Golay filtered data followed by the raw data. Then, the first derivative
of the filtered data achieves the next lowest slope, but it has to be noted that the
overall kappa score level of this data transformation is lower than for the derivatives
of the unfiltered data.



## Citations on this page
