---
title: "Support Vector Machine"
author: Darius Goergen
site: workflowr::wflow_site
output:
  workflowr::wflow_html:
    toc: true
editor_options:
  chunk_output_type: console
bibliography: library.bib
csl: elsevier-harvard.csl
link-citations: yes
---

```{r setup, include=FALSE, warning=FALSE, message=FALSE}
source("code/setup_website.R")
source("code/functions.R")
```

## Overview
Support Vector Machine (SVM) is a non-parametric classification method which was
initially designed for binary classification problems and was developed in 
its current form by @Boser:1992:TAO:130385.130401. A detailed overview of the 
SVM algorithm is found in @Burges1998. The principal idea behind SVM is to find 
an optimal hyperplane which separates two classes from each other by the largest
possible margin. The algorithm is optimized by iteratively maximizing this margin. 
Only the closest observations to the margin are considered. 
These specific observations are also called support-vectors. Multidimensional 
data can be processed by mapping the data into a higher dimensional feature 
space through a specified mapping function. This function is called kernel 
function, and mainly four different groups are used: linear, polynomial, radial 
and sigmoid functions [@Burges1998]. In this project only the radial basis 
function was used. Multi-class problems are addressed by calculating an optimal 
margin following the one class-against-all pattern and conducting a majority vote 
at the end of the calculations. SVM needs proper fine-tuning of its parameters. These are the 
regularization parameter `C` and the kernel width `γ` for radial basis functions. 
The regularization parameter is also called "penalty value", as it is a constant giving penalty to 
misclassified observations. The optimal values might change if different 
representations of the data are presented to the algorithm.


## Adding Noise

Different techniques of data pre-processing might emphasize different features of 
the patterns to be learned by an algorithm. To grasp this, different transformations
of the data were presented to the SVM algorithm. Additionally, the raw data signal
was jittered to test which transformation might prove beneficial in delivering 
a high classification accuracy even in the presence of noise. To test this, 
we define a function which adds noise to the raw data and returns a list with
the number of elements equal to the levels of noise applied.


```{r svm-addNoise}
addNoise = function(data, levels = c(0), category="class"){
  data.return = list()
  index = which(names(data) == category)
  for (n in levels){
    tmp = as.matrix(data[ , -index])
    if (n == 0){
      tmp = data
    }else{
      tmp = as.data.frame(jitter(tmp, n))
      tmp[category] = data[category]
    }
    data.return[[paste("noise", n, sep="")]] = tmp
  }
  return(data.return)
}

data = read.csv(file = paste0(ref, "reference_database.csv"), header = TRUE)
noisy_data = addNoise(data, levels = c(0,10,100,250,500), category = "class")

# indivitual elements can be selected by using [[ and refering to the index or the name
head(noisy_data[["noise100"]])[1:3,1:3]
```


## Data Pre-processing

In another function which uses the `noisy_data`-object as input,
specific data transformations are applied. These is normalization which centers
and scales the input data, different forms of the Savitzkiy-Golay filter [@Savitzky1964],
and first and second derivatives of the raw spectra. The function iterates through
the elements in the `noisy_data` object and returns each specified 
transformation in a list element below the noise level. The exemplary function below
applies the pre-processing for normalization, standard filtering and first 
derivative only. The implementation of the function used in the project
can be found [here](https://github.com/goergen95/polymeRID/blob/master/code/functions.R#12).

```{r svm-TrainigSet}
createTrainingSet = function(data, category = "class",
                             SGpara = list(p=3,w=11), lag = 15){
  
  data.return = list()
  for (noise in names(data)){
    
    tmp = as.data.frame(data[[noise]])
    classes = tmp[,category]
    tmp = tmp[!names(tmp) %in% category]
    
    # original data
    data.return[[noise]][["raw"]] = as.data.frame(data[[noise]])
    
    # normalised data
    data_norm = preprocess(tmp, type="norm")
    data_norm[category] = classes
    data.return[[noise]][["norm"]] = data_norm
    
    # SG-filtered data
    data_sg = preprocess(tmp, type="sg", SGpara = SGpara)
    data_sg[category] = classes
    data.return[[noise]][["sg"]] = data_sg
    
    # first derivative of original data
    data_rawd1 = preprocess(tmp, type="raw.d1", lag = lag)
    data_rawd1[category] = classes
    data.return[[noise]][["raw.d1"]] = data_rawd1
    
  }
  return(data.return)
}

# applying the function
test_dataset = createTrainingSet(noisy_data, category = "class")

# individual transformations at a certain noise level can be accessed with [[
head(test_dataset[["noise500"]][["raw.d1"]])[1:3,1:3]
```

## Dimensionality Reduction

The database of @Primpke2018 currently shows 1863 variables for each observation.
Most of these data points do not hold relevant information to distinguish between
different types of particles. To shorten the computation time, one can use 
dimensionality reduction techniques such as principal component analysis (PCA).
PCA has already been used to transform spectral data of microplastic in marine 
ecosystems [@Jung2018; @Lorenzo-Navarro2018]. PCA basically takes the input
data for a given number of observations and performs a orthogonal transformation
to  derive uncorrelated principal components from the possibly correlated variables.
Both redundancies in the data as well as the presence of noise can be accounted for this way. 
PCA has previously been successfully applied to FTIR-spectroscopy data [@Hori2003; @Nieuwoudt2004; @Mueller2013; @Ami2013; @Fu2014].
Simultaneously, the number of variables can be significantly reduced by applying
PCA and thus speeding up the training process. Below we apply a PCA to the
raw data as an example only. 

```{r PCA-raw , message=FALSE}
library(factoextra)
tmp = test_dataset[["noise0"]][["raw"]]
pca = prcomp(tmp[ ,-1864]) # omitting class variable
var_info = factoextra::get_eigenvalue(pca)
# setting a threshold of 99% explained variance
threshold = 99
thresInd = which(var_info$cumulative.variance.percent>=threshold)[1]
pca_data = pca$x[,1:thresInd]
```
We can use the index variable `thresInd` we just defined to take a look upon
all the principal components which explain 99% of the variance of the data.

```{r pca-table, echo=FALSE}
library(knitr)
knitr::kable(var_info[1:thresInd,])
```

We effectively reduced the number of variables from 1683 to 15 which still bears 
99% of the variance we find in the original data. When it comes
to machine learning, however, it is important to realize that this new data set is not fit
to be used in a training process. If we now randomly split the observations into
a training and testing set, we effectively mix up these two sets because information of 
the testing set has already influenced the outcome of the PCA. Therefore, the data set
needs to be split before applying the PCA. The analysis is done on the training data 
only and then the same orthogonal transformations is applied to the testing data.
This way it can be ensured that the test set is truly independent of the 
training process.

## Cross Validation

We apply a 10-fold cross-validation which is repeated five times. The following 
code takes a complete data set as input, applies a splitting function from the
`caret` package and then builds the PCA upon the the training set and finally applies
the same transformation to the testing set. Here, it is only applied for the raw data.
We also randomly split the data to a 50% training and a 50% testing set.

```{r PCA_loop}
folds = 10
repeats = 5
split_percentage = 0.5
threshold = 99
tmp = test_dataset[["noise0"]][["raw"]]

set.seed(42) # ensure reproducibility
fold_index = lapply(1:repeats, caret::createDataPartition, y=tmp$class,
                   times = folds, p = split_percentage)
fold_index = do.call(c, fold_index)

pcaData = list()
for (rep in 1:repeats){
  rep_index = fold_index[(rep*folds-folds+1):(rep*folds)] # jumps to the correct number of folds forward in each repeat
  
  pcadata_fold = lapply(1:folds,function(x){
    
    # splitting for current fold
    training = tmp[unlist(rep_index[x]),]
    validation = tmp[-unlist(rep_index[x]),]
    
    # keep response
    responseTrain = training$class
    responseVal = validation$class
    
    # apply PCA
    pca = prcomp(training[,1:1863])
    varInfo = factoextra::get_eigenvalue(pca)
    thresInd = which(varInfo$cumulative.variance.percent >= threshold)[1]
    pca_training = pca$x[ ,1:thresInd]
    pca_validation = predict(pca, validation)[ ,1:thresInd]
    
    training = as.data.frame(pca_training)
    training$response = responseTrain
    validation = as.data.frame(pca_validation)
    validation$response = responseVal
    foldtmp = list(training, validation)
    names(foldtmp) = c("training","validation")
    return(foldtmp)
  })
  names(pcadata_fold) = paste("fold", 1:folds, sep ="")
  pcaData[[paste0("repeat",rep)]] = pcadata_fold
}

```
We now have a list object with the number of elements equivalent to the repeats.
Below the level of repeats individual folds can be accessed.
There the splitted data set can be accessed by referring to `"training"`
and `"testing"`.
```{r PCA_access}
pcaData[["repeat5"]][["fold10"]][["training"]][1:3,1:3]
pcaData[["repeat5"]][["fold10"]][["validation"]][1:3,1:3]

summary(pcaData[["repeat5"]][["fold10"]][["training"]]$response)
summary(pcaData[["repeat5"]][["fold10"]][["validation"]]$response)
```

## Parameter Tuning

For the SVM algorithm a simple search pattern for the optimal configuration for
the regularization parameter `C` and the kernel width `γ` was implemented. We did
this by applying a search grid for the parameters and calculating each possible
combination. Note that due to limits in computation capacities we restricted
the search to 25 combinations only. The code below calculates a model for each
possible combination, evaluating its capacity to correctly classify the training
data and then chooses the optimal model to evaluate the validation data. 

```{r svm-paramter-tuning, eval = TRUE}
training = pcaData[["repeat1"]][["fold1"]][["training"]]
validation =pcaData[["repeat1"]][["fold1"]][["validation"]]
x_train = training[ ,1:ncol(training)-1]
y_train = training$response
x_test = validation[ ,1:ncol(validation)-1]
y_test = validation$response


tuneGrid = expand.grid(gamma =seq(0.1,1,0.2),cost = seq(1,5,1) )
accuracy = c()
models = list()
for ( i in 1:nrow(tuneGrid)){
  model = e1071::svm(x = x_train, y = y_train,
                    kernel = "radial",
                    gamma = tuneGrid$gamma[i],
                    cost = tuneGrid$cost[i])
  pred = predict(model, x_train)
  conf = caret::confusionMatrix(pred, y_train)
  accuracy = c(accuracy, conf$overall["Kappa"])
  models[[i]] = model
}

best_model = models[[which(accuracy == max(accuracy))[1]]]
prediction = predict(best_model, x_test)
confMat = caret::confusionMatrix(prediction, y_test)
print(confMat$table)
```

The process of splitting the data set into training and testing was automated by
putting the above code in a function which can be found [here](https://github.com/goergen95/polymeRID/blob/master/code/functions.R#239).
Finally, this function was applied to the different pre-procesesing levels, as discussed before.

```{r CV-apply, eval = FALSE}
source("code/functions.R")
wavenumbers = readRDS(paste0(ref,"wavenumbers.rds"))
# add noise to data
noisyData = addNoise(data,levels = c(0,10,100,250,500), category = "class")

# preprocessing
testDataset = createTrainingSet(noisyData, category = "class",
                                SGpara = list(p=3, w=11), lag=15,
                                type = c("raw", "norm", "sg", "sg.d1", "sg.d2",
                                  "sg.norm", "sg.norm.d1", "sg.norm.d2",
                                  "raw.d1", "raw.d2", "norm.d1", "norm.d2"))


types = names(testDataset[[1]])

levels = lapply(names(testDataset), function(x){
  rep(x, length(types))
})
levels = unlist(levels)

results = data.frame(level=levels, type = types, kappa = rep(0,length(levels)))

for (level in unique(levels)){
  for (type in types){

    print(paste0("Level: ",level," Type: ",type))
    tmpData = testDataset[[level]][[type]]
    tmpData[which(wavenumbers<=2420 & wavenumbers>=2200)] = 0 # setting C02 window to 0
    tmpModel = pcaCV(tmpData, folds = 10, repeats = 5, threshold = 99, metric = "Kappa", p=0.5, method="svm")
    saveRDS(tmpModel,file = paste0(output,"svm/model_",level,"_",type,"_",round(tmpModel[[1]],2),".rds"))
    results[which(results$level==level & results$type==type),"kappa"] = as.numeric(tmpModel[[1]])
    print(results)

  }
}
saveRDS(results, paste0(output,"svm/exploration.rds"))

```
```{r read-exploration, include = FALSE}
results = readRDS(paste0(output,"svm/exploration.rds"))
```


## Results 

The plot below shows the Kappa scores the algorithm achieved during training
for different representations of the data at increasing noise levels.

```{r plot-results, echo = FALSE, warning=FALSE, message=FALSE}
library(plotly)
results$level = as.character(results$level)
# let's visualize the accuracy results with noise levels on x-axis and kappa score on y-axis
test = ggplot(data=results)+
  geom_line(aes(y=kappa,group=type,x=level,color=type),size=1.5)+
  ylab("Kappa score")+
  xlab("Noise level")+
  scale_color_discrete(name="Type of\nPre-Processing")+
  theme_minimal()

t = ggplotly(test)
t
```

It can be observed that with higher noise level the Kappa score is reduced significantly.
All data transformations yield to very low accuracies when the noise increases.
In the absence of significant noise, however, the simple Savitzkiy-Golay filter,
the raw data, as well as the first and second order derivatives yield to a kappa
score of about 0.75. Looking at this more mathematically by calculating the average 
slopes of the Kappa scores reveals the transformations with the most stable results. 
Note that we only take the kappa score at noise level 0 and 10 to calculate the average slope,
since higher noise levels yielded to unsatisfactory accuracies.

```{r slope-calculation}
noise0 = results[results$level == "noise0", ]
noise10 = results[results$level == "noise10", ]
slopes = noise10$kappa - noise0$kappa 
types = unique(results$type)
df = data.frame(type = types, slope = slopes, row.names = NULL)
df = df[order(-slopes),]
```
```{r print-slope-table, echo = FALSE}
knitr::kable(df)
```

This confirms that the average decrease in Kappa score is the lowest for
the Savitzkiy-Golay filtered data followed by the raw data. Then, the first derivative
of the filtered data achieves the next lowest slope, but it has to be noted that the
overall kappa score level of this data transformation is lower than for the derivatives
of the unfiltered data.



## Citations on this page
